### 2016-06-13 paper/icfp16-cameraready/evaluation.tex
update benchmark numbers

\begin{figure*}[t]
\centering
\begin{minipage}{\linewidth}
\centering
\includegraphics[width=0.49\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
\includegraphics[width=0.49\linewidth]{distrib.png}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for 83\% of
  the programs in under one second, improving slightly to 87\% in under
  10 seconds. In both datasets we detect actual type errors about 82\%
  of the time, unbound variables or constructors 3--4\% of the time, and
  diverging loops 2--3\% of the time. For the remaining 11--12\% of the
  programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure*}

@@

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 31 for the
\ucsdbench\ dataset (35 for the \uwbench\ dataset) with a maximum of
2,745 (660 for \uwbench) and a median of 16 (also 16 for \uwbench).
%
The average number of jumps per trace is 7 (also 7 for \uwbench) with a
maximium of 353 (105 for \uwbench) and a median of 4 (also 4 for
\uwbench).
%
In both datasets 80\% or more traces have at most 10 jumps.
%

@@

\paragraph{Random Generation}
Random test generation has difficulty generating highly constrained
values, \eg red-black trees or a pair of equal integers. If the type
error is hidden behind a complex branch condition \nanomaly may not be
able to trigger it. Exhaustive testing and dynamic-symbolic execution
can address this short-coming by performing an exhaustive search for
inputs (\emph{resp}.\ paths through the program). As our experiments
show, however, novice programs do not appear to require more advanced
search techniques, likely because the novice programs tend to be simple.

\paragraph{Infinite Types}
Our implementation does check for infinite types inside \forcesym, but
there are some degenerate cases where it is unable to detect
them. Consider, the following buggy @replicate@
%
\begin{code}
  let rec replicate n x =
    if n <= 0 then []
              else replicate (n-1) [x]
\end{code}
%
This code produces a nested list (with @n@ levels of nesting) containing
a single copy of @x@, instead of a list with @n@ copies of @x@. \ocaml
detects a cyclic \hbox{@'a = 'a list@} constraint in the recursive call
and throws a type error, whereas \nanomaly happily % recurses @n@ times to
produces the nested list.  Strictly speaking, this function itself cannot
``go wrong'', the program would not get stuck until a \emph{client}
attempted to use the result expecting a flat list. But this is not very
satisfying as @replicate@ is clearly to blame. Furthermore, in our
experience, infinite-type errors are often difficult to %some of the more difficult ones to
debug (and to explain to novices), so better support for this scenario
would be useful.

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

\begin{figure*}[t]
\centering
\begin{minipage}{\linewidth}
\centering
\includegraphics[width=0.49\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
\includegraphics[width=0.49\linewidth]{distrib.png}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for 79--85\% of
  the programs in under one second, improving slightly to 87\% in under
  10 seconds. In both datasets we detect actual type errors about 82\%
  of the time, unbound variables or constructors 3--4\% of the time, and
  diverging loops 2--3\% of the time. For the remaining 11--12\% of the
  programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure*}

@@

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 31 for the
\ucsdbench\ dataset (35 for the \uwbench\ dataset) with a maximum of
2,745 (986 for \uwbench) and a median of 17 (also 17 for \uwbench).
%
The average number of jumps per trace is 7 (also 7 for \uwbench) with a
maximium of 353 (185 for \uwbench) and a median of 4 (also 4 for
\uwbench).
%
In both datasets 80\% or more traces have at most 10 jumps.
%

@@

\paragraph{Random Generation}
Random test generation has difficulty generating highly constrained
values, \eg\ red-black trees or a pair of equal integers. If the type
error is hidden behind a complex branch condition \nanomaly\ may not be
able to trigger it. Exhaustive testing and dynamic-symbolic execution
can address this short-coming by performing an exhaustive search for
inputs (\emph{resp}.\ paths through the program). As our experiments
show, however, novice programs do not appear to require more advanced
search techniques, likely because the novice programs tend to be simple.

\paragraph{Infinite Types}
Our implementation does check for infinite types inside \forcesym, but
there are some degenerate cases where it is unable to detect
them. Consider, the following buggy @replicate@
%
\begin{code}
  let rec replicate n x =
    if n <= 0 then []
              else replicate (n-1) [x]
\end{code}
%
This code produces a nested list (with @n@ levels of nesting) containing
a single copy of @x@, instead of a list with @n@ copies of @x@. \ocaml\
detects a cyclic \hbox{@'a = 'a list@} constraint in the recursive call
and throws a type error, whereas \nanomaly\ happily % recurses @n@ times to
produces the nested list.  Strictly speaking, this function itself cannot
``go wrong'', the program would not get stuck until a \emph{client}
attempted to use the result expecting a flat list. But this is not very
satisfying as @replicate@ is clearly to blame. Furthermore, in our
experience, infinite-type errors are often difficult to %some of the more difficult ones to
debug (and to explain to novices), so better support for this scenario
would be useful.

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
%!TEX root = main.tex

### 2016-06-13 paper/icfp16-cameraready/related.tex
update benchmark numbers

% \subparagraph{Localization}
The traditional Damas-Milner type inference
algorithm~\cite{Damas1982-uw} reports the first program location where a
type mismatch is discovered (subject to the traversal
strategy~\cite{Lee1998-ys}).
%
As a result the error can be reported far away from its
source~\cite{McAdam1998-ub} without enough information to guide the
user.
%
Type-error slicing~\cite{Haack2003-vc,Schilling2011-yf,Rahli2015-tt,Sagonas2013-bf,Gast2004-zd,Neubauer2003-xv}
%treats type inference as a constraint-satisfaction problem and
recognizes this flaw and instead produces a slice of the program
containing \emph{all} program locations that are connected to the type
error.
%
%% \citet{Neubauer2003-xv} present a decidable type system based on
%% discriminative sum types, in which all terms are typeable and type
%% derivations contain all type errors in a program. They then use the
%% typing derivation to slice out the parts of the expression related to
%% each error.
%
Though the program slice must contain the source of the error, it can
suffer from the opposite problem of providing \emph{too much}
information, motivating recent work in ranking the program locations.
%
Zhang~\etal~\cite{Zhang2014-lv,Zhang2015-yu} present an algorithm for
identifying the most likely culprit using Bayesian reasoning.
%
\citet{Pavlinovic2014-mr,Pavlinovic2015-kh} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.

>>>>

% \subparagraph{Localization}
The traditional Damas-Milner type inference
algorithm~\cite{Damas1982-uw} reports the first program location where a
type mismatch is discovered (subject to the traversal
strategy~\cite{Lee1998-ys}).
%
As a result the error can be reported far away from its
source~\cite{McAdam1998-ub} without enough information to guide the
user.
%
Type-error slicing~\cite{Haack2003-vc,Schilling2011-yf,Rahli2015-tt,Sagonas2013-bf,Gast2004-zd,Neubauer2003-xv}
%treats type inference as a constraint-satisfaction problem and
recognizes this flaw and instead produces a slice of the program
containing \emph{all} program locations that are connected to the type
error.
%
%% \citet{Neubauer2003-xv} present a decidable type system based on
%% discriminative sum types, in which all terms are typeable and type
%% derivations contain all type errors in a program. They then use the
%% typing derivation to slice out the parts of the expression related to
%% each error.
%
Though the program slice must contain the source of the error, it can
suffer from the opposite problem of providing \emph{too much}
information, motivating recent work in ranking the candidate locations.
%
Zhang~\etal~\cite{Zhang2014-lv,Zhang2015-yu} present an algorithm for
identifying the most likely culprit using Bayesian reasoning.
%
\citet{Pavlinovic2014-mr,Pavlinovic2015-kh} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.
